---
title: "Working iwth Dates and Times in R"
author: "KVP"
date: "9/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(reactable)
```

# Create Database
In this database we will have one column that has the date/times stored as a character and the other column storing the date/times as a numeric column. 

The numeric column is the seconds from 1970-01-01 00:00.00 UTC. When you are given a numeric date/time column be sure to check which software it was created with. I will be doing a follow-up post about working with excel numeric date/times.

```{r create_db, include = FALSE}
df <- data.frame(date_char = c("2021-09-15 09:15", "2020-10-05 20:20", "2021-11-03 13:00"),
                 date_numeric = as.numeric(as.POSIXlt(c("2021-09-15 09:15", "2020-10-05 20:20", 
                                                        "2021-11-03 13:00"))))
```

# Check the Type of Each Column
Now that we have our date/time columns created in our dataframe, we can check the type of each column by using the `class()` function. The `df$date_char` will be type `character` while `df$date_numeric` will be a `numeric` dataframe column.

```{r check_type}
# Prints a character type
class(df$date_char)

# Prints a numeric type
class(df$date_numeric)
```

# Parse Each Column
## Parse the Character Column
```{r parse_char}
df$date_char_parsed <- parse_date_time(df$date_char, "%Y-%m-%d %H:%M", tz = "America/New_York")
print(df$date_char_parsed)
```

## Parse the Numeric Column
```{r, parse_num}
df$date_numeric_parsed <- as.POSIXct(df$date_numeric, origin = "1970-01-01 00:00.00 UTC",
                                     tz = "America/New_York")
print(df$date_numeric_parsed)
```

# Summary
We now have our character and numeric date columns in the same format. We can now do calculations using `{lubridate}` or format the date for graphs or tables.

As a bonus, below is a table of all the different timezone names for your reference.

# Table of All the Timezone Names
```{r, timezone_names}
tzNames <- data.frame(names = OlsonNames())
tzNames %>%
  reactable(searchable = TRUE)
```

